---
title: '`Quiz 1` Exploratory Data Analysis'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
params:
  author: "Anderson H Uyekita"
  course: "Exploratory Data Analysis"
  course_url: "https://www.coursera.org/learn/exploratory-data-analysis"
  specialization: "Data Science: Foundations using R Specialization"
  specialization_url: "https://www.coursera.org/specializations/data-science-foundations-r"
  instructor: "Roger D Peng"
  week: "Week 1"
  course_start: !r lubridate::ymd("2022/05/25")
  course_finish: !r lubridate::ymd("2022/05/25")
  slug: 'Week%201'
always_allow_html: true
---

```{r setup, include=FALSE,message=FALSE,warning=FALSE,cache=TRUE}
# Defining document chunks options
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,message=FALSE,warning=FALSE)

# Loading required libraries
library(rmarkdown)
library(lubridate)
library(tidyverse)
library(magrittr)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
* `r if(!knitr::is_html_output()) {"&#x1F30E; Rpubs: [Interactive Document](https://rpubs.com/AndersonUyekita/quiz-1_exploratory-data-analysis)"}else{"&#x1F4E6; Github Repository: [Static Document](https://github.com/AndersonUyekita/getting-and-cleaning-data/blob/main/Week%201/quiz-1_exploratory-data-analysis.md){target='_blank' rel='noopener'}"}`

--------------------------------------------------------------------------------

```{r creating_folder,cache=TRUE,include=FALSE}
# Checking if the subfolder already exists.
if (!dir.exists("data")) {
    
    # Creating a subfolder to store the data.
    dir.create(path = "./data")
}
```

## Question 1

Which of the following is a principle of analytic graphics?


- [ ] Don't plot more than two variables at at time
- [x] Show multivariate data
- [ ] Show box plots (univariate summaries)
- [ ] Make judicious use of color in your scatterplots
- [ ] Only do what your tools allow you to do

> According to six principles stated during the video [Principles of Analytic Graphics](./slides/1_1_1_PrinciplesofAnalyticGraphics.pdf){target='_blank' rel='noopener'}:
>
> * Principle 1: Show comparisons
> * Principle 2: Show causality, mechanism, explanation
> * **Principle 3: Show multivariate data**
> * Principle 4: Integrate multiple modes of evidence
> * Principle 5: Describe and document the evidence
> * Principle 6: Content is king

## Question 2

What is the role of exploratory graphs in data analysis?

- [ ] Axes, legends, and other details are clean and exactly detailed.
- [x] They are typically made very quickly.
- [ ] They are made for formal presentations.
- [ ] Only a few are constructed.

> According to the Summary of [Exploratory Graphs](./slides/1_1_2_ExploratoryGraphs.pdf){target='_blank' rel='noopener'} document.
>
> * **Exploratory plots are "quick and dirty"**
> * Let you summarize the data (usually graphically) and highlight any broad features
> * Explore basic questions and hypotheses (and perhaps rule them out)
> * Suggest modeling strategies for the "next step"

## Question 3

Which of the following is true about the base plotting system?

- [x] Plots are created and annotated with separate functions
- [ ] Margins and spacings are adjusted automatically depending on the type of plot and the data
- [ ] Plots are typically created with a single function call
- [ ] The system is most useful for conditioning plots

> According to the Summary of [The Base Plotting System](./slides/1_2_2_PlottingBase.pdf){target='_blank' rel='noopener'} in R document.
>
> * Plots in the base plotting system are created by calling successive R functions to "build up" a plot
> * **Plotting occurs in two stages:**
>   * **Creation of a plot**
>   * **Annotation of a plot (adding lines, points, text, legends)**
> * The base plotting system is very flexible and offers a high degree of control over plotting

## Question 4

Which of the following is an example of a valid graphics device in R?

- [ ] The keyboard
- [ ] A socket connection
- [ ] A Microsoft Word document
- [ ] A file folder
- [x] A PNG file

> Based on slides 6 and 7 from [Graphics Devices in R](./slides/1_3_GraphicsDevicesinR.pdf){target='_blank' rel='noopener'} document.
> 
> * Vector formats: `pdf`, `svg`, `win.metafile`, and `postscript`, and;
> * Bitmap formats: `png`, `jpeg`, `tiff`, and `bmp`.


## Question 5

Which of the following is an example of a vector graphics device in R?

- [x] Postscript
- [ ] TIFF
- [ ] JPEG
- [ ] GIF
- [ ] PNG

> Based on slide 6 from [Graphics Devices in R](./slides/1_3_GraphicsDevicesinR.pdf){target='_blank' rel='noopener'} document.
> 
> * Vector formats: `pdf`, `svg`, `win.metafile`, and `postscript`, and;

## Question 6

Bitmapped file formats can be most useful for 

- [x] Scatterplots with many many points
- [ ] Plots that are not scaled to a specific resolution
- [ ] Plots that require animation or interactivity
- [ ] Plots that may need to be resized

> According to the Summary of [Graphics Devices in R](./slides/1_3_GraphicsDevicesinR.pdf){target='_blank' rel='noopener'} in R document.
>
> * Plots must be created on a graphics device
> * The default graphics device is almost always the screen device, which is most useful for exploratory analysis
> * File devices are useful for creating plots that can be included in other documents or sent to other people
> * For file devices, there are vector and bitmap formats
>   * Vector formats are good for line drawings and plots with solid colors using a modest number of points
>   * **Bitmap formats are good for plots with a large number of points, natural scenes or webbased plots**

## Question 7

Which of the following functions is typically used to add elements to a plot in the base graphics system?

- [x] text()
- [ ] plot()
- [ ] boxplot()
- [ ] hist()

> Based on slide 13 from [The Base Plotting System in R](./slides/1_2_2_PlottingBase.pdf){target='_blank' rel='noopener'} document.
> 
> * plot: make a scatterplot, or other type of plot depending on the class of the object being plotted
> * lines: add lines to a plot, given a vector x values and a corresponding vector of y values (or a 2-
> * column matrix); this function just connects the dots
> * points: add points to a plot
> * **text: add text labels to a plot using specified x, y coordinates**
> * title: add annotations to x, y axis labels, title, subtitle, outer margin
> * mtext: add arbitrary text to the margins (inner or outer) of the plot
> * axis: adding axis ticks/labels

## Question 8

Which function opens the screen graphics device for the Mac?

- [x] quartz()
- [ ] png()
- [ ] bitmap()
- [ ] pdf()

> Based on slide 2 from [Graphics Devices in R](./slides/1_3_GraphicsDevicesinR.pdf){target='_blank' rel='noopener'} document.
>
> * A graphics device is something where you can make a plot appear
>   * A window on your computer (screen device)
>   * A PDF file (file device)
>   * A PNG or JPEG file (file device)
>   * A scalable vector graphics (SVG) file (file device)
> * When you make a plot in R, it has to be "sent" to a specific graphics device
> * The most common place for a plot to be "sent" is the screen device
>   * **On a Mac the screen device is launched with the quartz()**
>   * On Windows the screen device is launched with windows()
>   * On Unix/Linux the screen device is launched with x11()

## Question 9

What does the 'pch' option to par() control?

- [ ] the orientation of the axis labels on the plot
- [ ] the size of the plotting symbol in a scatterplot
- [x] the plotting symbol/character in the base graphics system
- [ ] the line width in the base graphics system

> Based on slide 9 from [The Base Plotting System in R](./slides/1_2_2_PlottingBase.pdf){target='_blank' rel='noopener'} document.
>
> * **pch: the plotting symbol (default is open circle)**
> * lty: the line type (default is solid line), can be dashed, dotted, etc.
> * lwd: the line width, specified as an integer multiple
> * col: the plotting color, specified as a number, string, or hex code; the colors() function gives
> * you a vector of colors by name
> * xlab: character string for the x-axis label
> * ylab: character string for the y-axis label

## Question 10

If I want to save a plot to a PDF file, which of the following is a correct way of doing that?

- [ ] Construct the plot on the PNG device with png(), then copy it to a PDF with dev.copy2pdf().
- [x] Construct the plot on the screen device and then copy it to a PDF file with dev.copy2pdf()
- [ ] Open the screen device with quartz(), construct the plot, and then close the device with dev.off().
- [ ] Open the PostScript device with postscript(), construct the plot, then close the device with dev.off().

> Based on slide 9 from [Graphics Devices in R](./slides/1_3_GraphicsDevicesinR.pdf){target='_blank' rel='noopener'} document.
>
> Copying a plot to another device can be useful because some plots require a lot of code and it can be a pain to type all that in again for a different device.
>
> * dev.copy: copy a plot from one device to another
> * **dev.copy2pdf: specifically copy a plot to a PDF file**

```{r,include=FALSE}
markdown::renderMarkdown(file = basename(path = rstudioapi::getSourceEditorContext()$path))
```