---
title: '`Week 1` Exploratory Data Analysis'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
params:
  author: "Anderson H Uyekita"
  course: "Exploratory Data Analysis"
  course_url: "https://www.coursera.org/learn/exploratory-data-analysis"
  specialization: "Data Science: Foundations using R Specialization"
  specialization_url: "https://www.coursera.org/specializations/data-science-foundations-r"
  instructor: "Roger D Peng"
  week: "Week 1"
  course_start: !r lubridate::ymd("2022/05/25")
  course_finish: !r lubridate::ymd("2022/05/25")
  slug: 'Week%201'
---

```{r setup, include=FALSE,message=FALSE,warning=FALSE,cache=TRUE}
# Defining document chunks options
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,message=FALSE,warning=FALSE)

# Loading required libraries
library(rmarkdown)
library(lubridate)
```

* `r if(!knitr::is_html_output()) {sprintf(fmt = "&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
* `r sprintf(fmt = "&#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
* `r sprintf(fmt = "&#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    * `r sprintf(fmt = "&#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
* `r sprintf(fmt = "&#x1F4C6; %s", params$week)`
    * `r sprintf(fmt = "&#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    * `r sprintf(fmt = "&#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`

--------------------------------------------------------------------------------

#### Assignments & Deliverables

* &#x1F4BB; Swirl
    * Principles of Analytic Graphs
    * Exploratory Graphs
    * Graphics Devices in R
    * Plotting Systems
    * Base Plotting System
* [&#x1F4DD; Quiz 1](./quiz-1_exploratory-data-analysis.md)
* [&#x1F680; Course Project 1 Repository](https://github.com/AndersonUyekita/exploratory-data-analysis_course-project-1)

--------------------------------------------------------------------------------

#### Slides

* Lesson 1: Graphs [](){id="graphs"}
    * [Principles of Analytic Graphics](./slides/1_1_1_PrinciplesofAnalyticGraphics.pdf)
    * [Exploratory Graphs](./slides/1_1_2_ExploratoryGraphs.pdf)
* Lesson 2: Plotting [](){id="plotting"}
    * [Plotting Systems in R](./slides/1_2_1_Plotting Systems in R.pdf)
    * [Base Plotting System](./slides/1_2_2_PlottingBase.pdf)
* Lesson 3: Graphics Devices [](){id="graphics_devices"}
    * [Graphics Devices in R](./slides/1_3_GraphicsDevicesinR.pdf)

--------------------------------------------------------------------------------

#### Description

> This week covers the basics of analytic graphics and the base plotting system in R. We've also included some background material to help you install R if you haven't done so already.

--------------------------------------------------------------------------------

## Class Notes

### [<kbd>Lesson 1</kbd>](#graphs) Principles of Analytic Graphics

**Summary**

* Principle 1: Show comparisons

    > * Evidence for a hypothesis is always relative to another competing hypothesis.
    > * Always ask "Compared to What?"

* Principle 2: Show causality, mechanism, explanation

    > * What is your causal framework for thinking about a question?

* Principle 3: Show multivariate data

    > * Multivariate = more than 2 variables
    > * The real world is multivariate
    > * Need to "escape flatland"

* Principle 4: Integrate multiple modes of evidence

    > * Completely integrate words, numbers, images, diagrams
    > * Data graphics should make use of many modes of data presentation
    > * Don't let the tool drive the analysis

* Principle 5: Describe and document the evidence

    > * A data graphic should tell a complete story that is credible

* Principle 6: Content is king

    > * Analytical presentations ultimately stand or fall depending on the quality, relevance, and
integrity of their content

### [<kbd>Lesson 1</kbd>](#graphs) Exploratory Graphs

>* To understand data properties
>* To find patterns in data
>* To suggest modeling strategies
>* To "debug" analyses
>* To communicate results

The exploratory graphs are really about the **first four** things on this list.

Characteristics of exploratory graphs:

>* They are made quickly
>* A large number are made
>* The goal is for personal understanding
>* Axes/legends are generally cleaned up (later)
>* Color/size are primarily used for information

**Summary**

>* Exploratory plots are "quick and dirty"
>* Let you summarize the data (usually graphically) and highlight any broad features
>* Explore basic questions and hypotheses (and perhaps rule them out)
>* Suggest modeling strategies for the "next step"

### [<kbd>Lesson 2</kbd>](#plotting) Plotting Systems in R

**Summary**

There are three basic ways to plot a graph in R.

1. Base Plotting System: "artist's palette" model
    
    > * Convenient, mirrors how we think of building plots and analyzing data
    > * Can’t go back once plot has started (i.e. to adjust margins); need to plan in advance
    > * Difficult to "translate" to others once a new plot has been created (no graphical "language")
    > * Plot is just a series of R commands

2. Lattice: Entire plot specified by one function; conditioning

    > * Plots are created with a single function call (xyplot, bwplot, etc.)
    > * Most useful for conditioning types of plots: Looking at how y changes with x across levels of z
    > * Things like margins/spacing set automatically because entire plot is specified at once
    > * Good for putting many many plots on a screen
    > * Sometimes awkward to specify an entire plot in a single function call
    > * Annotation in plot is not especially intuitive
    > * Use of panel functions and subscripts difficult to wield and requires intense preparation
    > * Cannot "add" to the plot once it is created

3. ggplot2: Mixes elements of Base and Lattice

    > * Splits the difference between base and lattice in a number of ways
    > * Automatically deals with spacings, text, titles but also allows you to annotate by "adding" to a plot
    > * Superficial similarity to lattice but generally easier/more intuitive to use
    > * Default mode makes many choices for you (but you can still customize to your heart's desire)


### [<kbd>Lesson 2</kbd>](#plotting) Base Plotting System in R

Some Important Base Graphics Parameters Some Important Base Graphics Parameters

* `pch`: the plotting symbol (default is open circle)
* `lty`: the line type (default is solid line), can be dashed, dotted, etc.
* `lwd`: the line width, specified as an integer multiple
* `col`: the plotting color, specified as a number, string, or hex code; the colors() function gives
* `you` a vector of colors by name
* `xlab`: character string for the x-axis label
* `ylab`: character string for the y-axis label

The par() function is used to specify global graphics parameters

* `las`: the orientation of the axis labels on the plot
* `bg`: the background color
* `mar`: the margin size
* `oma`: the outer margin size (default is 0 for all sides)
* `mfrow`: number of plots per row, column (plots are filled row-wise)
* `mfcol`: number of plots per row, column (plots are filled column-wise)

```{r}
# Loading toy datasets.
library(datasets)

# Defining global parameters
par(mfrow = c(1, 3),      # c("number of rows", "number of columns")
    mar = c(4, 4, 2, 1),  # c(bottom, left, top, right)
    oma = c(0, 0, 2, 0))  # c(bottom, left, top, right)

# Simple plot as example.
with(airquality, {
    plot(Wind, Ozone, main = "Ozone and Wind")
    plot(Solar.R, Ozone, main = "Ozone and Solar Radiation")
    plot(Temp, Ozone, main = "Ozone and Temperature")
    mtext("Ozone and Weather in New York City", outer = TRUE)
})
```

**Summary**

> * Plots in the base plotting system are created by calling successive R functions to "build up" a plot
> * Plotting occurs in two stages:
>   * Creation of a plot
>   * Annotation of a plot (adding lines, points, text, legends)
> * The base plotting system is very flexible and offers a high degree of control over plotting

### [<kbd>Lesson 3</kbd>](#graphics_devices) Graphics Devices in R

Display a plot in the R viewer panel.

```{r,include=TRUE}
# Step 1 - Loading datasets to use mtcars.
library(datasets)

# Step 2 - Creating a Graph
with(mtcars, plot(wt, mpg)) # This is an alternative to:
                            # plot(mtcars$wt, mtcars$mpg)

# Step 3 - Adding notations
title(main = "Miles Per Gallon vs. Cars' Weight")
```

Exporting the graph into an external file called `myplot.pdf`, and storing in the data folder.

```{r,message=FALSE,warning=FALSE}
# Open PDF device to create the file and store it in the data folder.
grDevices::pdf(file = "./data/myplot.pdf")

# Previous graph example here.
with(mtcars, plot(wt, mpg))
title(main = "Miles Per Gallon vs. Cars' Weight")

# Closing the device
dev.off()
```

There are two basic types of file devices: vector and bitmap devices

* Vector formats: `pdf`, `svg`, `win.metafile`, and `postscript`, and;
* Bitmap formats: `png`, `jpeg`, `tiff`, and `bmp`.

**Summary**

> * Plots must be created on a graphics device
> * The default graphics device is almost always the screen device, which is most useful for exploratory analysis
> * File devices are useful for creating plots that can be included in other documents or sent to other people
> * For file devices, there are vector and bitmap formats
>   * Vector formats are good for line drawings and plots with solid colors using a modest number of points
>   * Bitmap formats are good for plots with a large number of points, natural scenes or webbased plots