---
title: '`Quiz 2` Exploratory Data Analysis'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
  github_document: default
params:
  author: "Anderson H Uyekita"
  course: "Exploratory Data Analysis"
  course_url: "https://www.coursera.org/learn/exploratory-data-analysis"
  specialization: "Data Science: Foundations using R Specialization"
  specialization_url: "https://www.coursera.org/specializations/data-science-foundations-r"
  instructor: "Roger D Peng"
  week: "Week 2"
  course_start: !r lubridate::ymd("2022/06/07")
  course_finish: !r lubridate::ymd("2022/06/08")
  slug_relative: 'https://github.com/AndersonUyekita/exploratory-data-analysis/'
  slug_chapter: 'Week%202/'
  lesson_1: '2_1_plotting-lattice.pdf'
  lesson_2: '2_2_ggplot-2.pdf'
always_allow_html: true
---

```{r setup, include=FALSE,message=FALSE,warning=FALSE,cache=TRUE}
# Defining document chunks options
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,message=FALSE,warning=FALSE)

# Loading required libraries
library(rmarkdown)
library(lubridate)
library(tidyverse)
library(magrittr)
library(lattice)
library(ggplot2)
library(ggplot2movies)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
* `r if(!knitr::is_html_output()) {"&#x1F30E; Rpubs: [Interactive Document](https://rpubs.com/AndersonUyekita/quiz-2_exploratory-data-analysis)"}else{"&#x1F4E6; Github Repository: [Static Document](https://github.com/AndersonUyekita/getting-and-cleaning-data/blob/main/Week%201/quiz-2_exploratory-data-analysis.md){target='_blank' rel='noopener'}"}`

--------------------------------------------------------------------------------

```{r creating_folder,cache=TRUE,include=FALSE}
# Checking if the subfolder already exists.
if (!dir.exists("data")) {
    
    # Creating a subfolder to store the data.
    dir.create(path = "./data")
}
```

## Question 1

Under the lattice graphics system, what do the primary plotting functions like `xyplot()` and `bwplot()` return?

- [ ] nothing; only a plot is made
- [x] an object of class "trellis"
- [ ] an object of class "plot"
- [ ] an object of class "lattice"

> According to slide 7 from [the Lattice Plotting System in R][2_1_plotting-lattice.pdf] document:
>
> * Lattice graphics functions return an object of the class trellis.

## Question 2

What is produced by the following code?

```
library(nlme)
library(lattice)
xyplot(weight ~ Time | Diet, BodyWeight)
```

- [x] A set of 3 panels showing the relationship between weight and time for each diet.
- [ ] A set of 3 panels showing the relationship between weight and time for each rat.
- [ ] A set of 11 panels showing the relationship between weight and diet for each time.
- [ ] A set of 16 panels showing the relationship between weight and time for each rat.

First, we need to investigate the `nlme` package and what is inside of `BodyWeight` dataset.

```{r,message=FALSE,warning=FALSE}
library(nlme)
library(lattice)

# First rows of BodyWeight dataset.
head(nlme::BodyWeight)

# Counting the number of Diet factors
length(unique(nlme::BodyWeight$Diet))
```

We already know the graphic should present `r length(unique(nlme::BodyWeight$Diet))` facets.

Also, according to the Lattice Graphic notation, `weight` will be on the y-axis and `Time` on the x-axis.

```{r}
# Printing the Lattice Graphic to prove it.
xyplot(weight ~ Time | Diet, BodyWeight)
```

**The graphic will display three facets presenting how each `Diet` will affect the Weight over `Time`.**

## Question 3

Annotation of plots in any plotting system involves adding points, lines, or text to the plot, in addition to customizing axis labels or adding titles. Different plotting systems have different sets of functions for annotating plots in this way.

Which of the following functions can be used to annotate the panels in a multi-panel lattice plot?

- [ ] points()
- [ ] lines()
- [x] panel.lmline()
- [ ] axis()
- [x] llines()

Based on slides 11 and 12 from [The Lattice Plotting System in R][2_1_plotting-lattice.pdf] document, I found a way to add an annotation in Lattice using the `panel.xyz()`, where `xyz()` could be abline lmline or other function.

The example on slide 12 uses the `panel.lmline()` annotation.

To better understand `llines()`, I have used the `?llines`.

## Question 4

The following code does NOT result in a plot appearing on the screen device.

```
library(lattice)
library(datasets)
data(airquality)
p <- xyplot(Ozone ~ Wind | factor(Month), data = airquality)
```

Which of the following is an explanation for why no plot appears?

- [ ] The xyplot() function, by default, sends plots to the PDF device.
- [ ] The variables being plotted are not found in that dataset.
- [ ] There is a syntax error in the call to xyplot().
- [x] The object 'p' has not yet been printed with the appropriate print method.

The graphic is OK unless it lacks a printing command to show it. The auto-printing will only work if the plot is not assigned to a variable, which is the current case.

## Question 5

In the lattice system, which of the following functions can be used to finely control the appearance of all lattice plots?

- [x]trellis.par.set()
- [ ]splom()
- [ ]par()
- [ ]print.trellis()

`?trellis.par.set`.

> **Description**
>
> Functions used to query, display and modify graphical parameters for fine control of Trellis displays. Modifications are made to the settings for the currently active device only.

## Question 6

What is ggplot2 an implementation of?

- [ ] the base plotting system in R
- [ ] the S language originally developed by Bell Labs
- [ ] a 3D visualization system
- [x] the Grammar of Graphics developed by Leland Wilkinson

Found in slide 2 of [ggplot2][2_2_ggplot-2.pdf] document.

> An implementation of the Grammar of Graphics by Leland Wilkinson

## Question 7

Load the `airquality' dataset form the datasets package in R

```
library(datasets)
data(airquality)
```

I am interested in examining how the relationship between ozone and wind speed varies across each month. What would be the appropriate code to visualize that using ggplot2? 

- [ ] `qplot(Wind, Ozone, data = airquality, facets = . ~ factor(Month))`
- [ ] `qplot(Wind, Ozone, data = airquality)`
- [ ] `qplot(Wind, Ozone, data = airquality, geom = "smooth")`
- [x] `airquality = transform(airquality, Month = factor(Month)); qplot(Wind, Ozone, data = airquality, facets = .~Month)`

The `transform()` function explicitly ensures the Month variable will be a factor. Although, the `. ~factor(Month)` is more neatly for me.

## Question 8

What is a `geom` in the ggplot2 system?

- [ ] a statistical transformation
- [x] a plotting object like point, line, or other shape
- [ ] a method for making conditioning plots
- [ ] a method for mapping data to attributes like color and size

According to the slide 24 of [Ggplot2][2_2_ggplot-2.pdf] Document.

> * `geoms`: geometric objects like points, lines, shapes.

## Question 9

When I run the following code I get an error:

```
library(ggplot2)
library(ggplot2movies)
g <- ggplot(movies, aes(votes, rating))
print(g)
```

I was expecting a scatterplot of 'votes' and 'rating' to appear. What's the problem?

- [x] ggplot does not yet know what type of layer to add to the plot.
- [ ] There is a syntax error in the call to ggplot.
- [ ] The object 'g' does not have a print method.
- [ ] The dataset is too large and hence cannot be plotted to the screen.

It is necessary to install the `ggplot2movies` package. The problem relies on the lack of geom method to select the way graphic will display the data.

## Question 10

The following code creates a scatterplot of 'votes' and 'rating' from the movies dataset in the ggplot2 package. After loading the ggplot2 package with the library() function, I can run

```
qplot(votes, rating, data = movies)
```

How can I modify the the code above to add a smoother to the scatterplot?

- [ ] qplot(votes, rating, data = movies) + stats_smooth("loess")
- [ ] qplot(votes, rating, data = movies, smooth = "loess")
- [x] qplot(votes, rating, data = movies) + geom_smooth()
- [ ] qplot(votes, rating, data = movies, panel = panel.loess)

There are two ways to do it:

Option 1 - Example on slide 9 from [Gpplot2][2_2_ggplot-2.pdf] document

```{r,cache=TRUE,message=FALSE,warning=FALSE}
# Adding geom argument on qplot function:
qplot(votes, rating, data = movies, geom = c("point","smooth"))
```

Option 2 - Example on slide 19 from [Gpplot2][2_2_ggplot-2.pdf] document
```{r,cache=TRUE,message=FALSE,warning=FALSE}
# Adding geom_smooth to qplot function:
qplot(votes, rating, data = movies) + geom_smooth()
```

```{r,include=FALSE}
markdown::renderMarkdown(file = basename(path = rstudioapi::getSourceEditorContext()$path))
```




```{r,include=FALSE}
# Github Document
if(!knitr::is_html_output()) {
    "Eu posso trabalhar com o endereço relativo"
    ref_1 <- "./slides/2_1_plotting-lattice.pdf"
    ref_2 <- "./slides/2_2_ggplot-2.pdf"

# HTML document
} else {
    "Eu preciso do URL completo"
    ref_1 <- paste0(params$slug_relative,"tree/main/",params$slug_chapter,"slides/2_1_plotting-lattice.pdf")
    ref_2 <- paste0(params$slug_relative,"tree/main/",params$slug_chapter,"slides/2_2_ggplot-2.pdf")
}
```

<!-- Inner references -->

[2_1_plotting-lattice.pdf]: `r ref_1`
[2_2_ggplot-2.pdf]: `r ref_2`
