---
title: '`Quiz 2` Reproducible Research'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  html_document:
    self_contained: yes
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
      smooth_scroll: true
  github_document: default
params:
  author: "Anderson H Uyekita"
  course: "Reproducible Research"
  course_url: "https://www.coursera.org/learn/reproducible-research"
  specialization: "Data Science: Foundations using R Specialization"
  specialization_url: "https://www.coursera.org/specializations/data-science-foundations-r"
  instructor: "Roger D Peng"
  course_start: !r lubridate::ymd("2022/06/23")
  course_finish: !r lubridate::ymd("2022/06/23")
  week: "Week 2"
  slug: 'Week%202'
always_allow_html: true
---

```{r setup, include=FALSE,message=FALSE,warning=FALSE,cache=TRUE}
# Defining document chunks options
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,message=FALSE,warning=FALSE)

# Loading required libraries
library(rmarkdown)
library(lubridate)
library(tidyverse)
library(magrittr)
```

`r if(!knitr::is_html_output()) {sprintf(fmt = "* &#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
`r sprintf(fmt = "* &#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
`r sprintf(fmt = "* &#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    `r sprintf(fmt = "* &#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
`r sprintf(fmt = "* &#x1F4C6; %s", params$week)`
    `r sprintf(fmt = "* &#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    `r sprintf(fmt = "* &#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`
* `r if(!knitr::is_html_output()) {"&#x1F30E; Rpubs: [Interactive Document](https://rpubs.com/AndersonUyekita/quiz-2_reproducible-research)"}else{"&#x1F4E6; Github Repository: [Static Document](https://github.com/AndersonUyekita/reproducible-research/blob/main/Week%202/quiz-2_reproducible-research.md){target='_blank' rel='noopener'}"}`

--------------------------------------------------------------------------------

```{r creating_folder,cache=TRUE,include=FALSE}
# Checking if the subfolder already exists.
if (!dir.exists("data")) {
    
    # Creating a subfolder to store the data.
    dir.create(path = "./data")
}
```

## Question 1

Who created Markdown?

- [ ] Yihui Xie
- [x] John Gruber
- [ ] Robert Gentleman
- [ ] Hadley Wickham

**Answer**

What is Markdown? What is Markdown?

- Created by John Gruber and Aaron Swartz
- A simplified version of "markup" languages
- Allows one to focus on writing as opposed to formatting
- Simple/minimal intuitive formatting elements
- Easily converted to valid HTML (and other formats) using existing tools
- Complete information is available at http://daringfireball.net/projects/markdown/
- Some background information at http://daringfireball.net/2004/03/dive_into_markdown

## Question 2

When writing a document in R Markdown, how do you denote the beginning of an R code chunk?

- [ ] \<rcode\>
- [ ] '''
- [x] '''{r}
- [ ] \<code\>

## Question 3

When using knitr, how do you indicate the height and width of a plot created in a code chunk?

- [ ] Set the 'height' and 'width' options for the code chunk
- [ ] Set the 'size' and 'scale' options for the code chunk
- [x] Set the 'fig.height' and 'fig.width' options for the code chunk
- [ ] Set the 'dpi' option for the code chunk

**Answer**

According to the Yihui Xie on its [webpage](https://yihui.org/knitr/options/#plots):

>fig.width, fig.height: (both are 7; numeric) Width and height of the plot (in inches), to be used in the graphics device.

## Question 4

With some code chunks, we may not want the output generated by the chunk to be rendered into HTML but would prefer to print the output verbatim. How can we specify this preference for a given code chunk?

- [ ] Set the option message = FALSE
- [ ] Set the option highlight = TRUE
- [x] Set the option results = "asis"
- [ ] Set the option tidy = FALSE

**Answer**

```{r mtcars-plots, results='asis'}
# Example.
cat("**Hello** World!")
```

## Question 5

When using knitr and R Markdown and producing output in HTML, why should you never edit the resulting HTML file?

- [x] Every time you knit() the R Markdown file, the HTML file will be overwritten
- [ ] The Markdown file generated by knitr is the appropriate file to edit.
- [ ] Editing the HTML file requires knowledge of a separate markup language.
- [ ] The HTML file is not a text file.

**Answer**

Processing of knitr Documents (what happens under the hood):

* You write the RMarkdown document (`.Rmd`)
* knitr produces a Markdown document (`.md`)
* knitr converts the Markdown document into HTML (by default)
* `.Rmd` -> `.md` -> `.html`
    * **This process will overwrite the `.md` and `.html` files.**
* You should NOT edit (or save) the `.md` or `.html` documents until you are finished
