---
title: '`Week 2` Exploratory Data Analysis'
author: '`r if(knitr::is_html_output()) {"&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Anderson H Uyekita"} else {NULL}`'
output:
  github_document: default
params:
  author: "Anderson H Uyekita"
  course: "Exploratory Data Analysis"
  course_url: "https://www.coursera.org/learn/exploratory-data-analysis"
  specialization: "Data Science: Foundations using R Specialization"
  specialization_url: "https://www.coursera.org/specializations/data-science-foundations-r"
  instructor: "Roger D Peng"
  week: "Week 2"
  course_start: !r lubridate::ymd("2022/06/07")
  course_finish: !r lubridate::ymd("2022/06/08")
  slug: 'Week%202'
---

```{r setup, include=FALSE,message=FALSE,warning=FALSE,cache=TRUE}
# Defining document chunks options
knitr::opts_chunk$set(echo = TRUE,cache=TRUE,message=FALSE,warning=FALSE)

# Loading required libraries
library(rmarkdown)
library(lubridate)
```

* `r if(!knitr::is_html_output()) {sprintf(fmt = "&#x1f468;&#x1F3FB;&#x200d;&#x1f4bb; Author: %s", params$author)}`
* `r sprintf(fmt = "&#x1f4da; Specialization: [%s](%s){target='_blank' rel='noopener'}", params$specialization, params$specialization_url)`
* `r sprintf(fmt = "&#x1f4d6; Course: [%s](%s){target='_blank' rel='noopener'}", params$course, params$course_url)`
    * `r sprintf(fmt = "&#x1F9D1;&#x200d;&#x1F3EB; Instructor: %s", params$instructor)`
* `r sprintf(fmt = "&#x1F4C6; %s", params$week)`
    * `r sprintf(fmt = "&#x1F6A6; Start: %s", format(params$course_start, "%A, %d %B %Y"))`
    * `r sprintf(fmt = "&#x1F3C1; Finish: %s", format(params$course_finish, "%A, %d %B %Y"))`

--------------------------------------------------------------------------------

#### Assignments & Deliverables

* &#x1F4BB; Swirl
    * Lattice Plotting System
    * Working with Colors
    * GGPlot2 Part1
    * GGPlot2 Part2
    * GGPlot2 Extras
* [&#x1F4DD; Quiz 2](./quiz-2_exploratory-data-analysis.md)

--------------------------------------------------------------------------------

#### Slides

* Lesson 1: Lattice Plotting System [](){id="lesson-1"}
    * [Lattice Plotting System](./slides/2_1_plotting-lattice.pdf)
* Lesson 2: ggplot2 [](){id="lesson-2"}
    * [ggplot2](./slides/2_2_ggplot-2.pdf)

--------------------------------------------------------------------------------

#### Description

> Welcome to Week 2 of Exploratory Data Analysis. This week covers some of the more advanced graphing systems available in R: the Lattice system and the ggplot2 system. While the base graphics system provides many important tools for visualizing data, it was part of the original R system and lacks many features that may be desirable in a plotting system, particularly when visualizing high dimensional data. The Lattice and ggplot2 systems also simplify the laying out of plots making it a much less tedious process.

--------------------------------------------------------------------------------

## Class Notes

### [<kbd>Lesson 1</kbd>](#lesson-1) Lattice Plotting System in R

**The Lattice Plotting System**

The lattice plotting system is implemented using `lattice` and `grid` packages.

> * lattice: contains code for producing Trellis graphics, which are independent of the “base” graphics system; includes functions like `xyplot`, `bwplot`, `levelplot`
> * grid: implements a different graphing system independent of the “base” system; the lattice package builds on top of grid
> * The lattice plotting system does not have a "two-phase" aspect with separate plotting and annotation like in base plotting
> * All plotting/annotation is done at once with a single function call

**Lattice Functions**

> * `xyplot`: this is the main function for creating scatterplots
> * `bwplot`: box-and-whiskers plots (“boxplots”)
> * `histogram`: histograms
> * `stripplot`: like a boxplot but with actual points
> * `dotplot`: plot dots on "violin strings"
> * `splom`: scatterplot matrix; like `pairs` in base plotting system
> * `levelplot`, contourplot: for plotting "image" data

Lattice functions generally take a formula for their first argument, usually of the form

```
xyplot(y ~ x | f * g, data)
```
We use the formula notation here, hence the `~` on the left of the `~` is the y-axis variable, on the right is the x-axis variable.

```{r}
library(datasets)
library(lattice)

# Scatter plot example.
xyplot(x = mpg ~ wt, data = mtcars)
```

`f` and `g` are conditioning variables — they are optional, the `*` indicates an interaction between two variables.

```{r}
# Scatter plot example with facets.
xyplot(x = mpg ~ wt | factor(cyl), data = mtcars)
```

The second argument is the data frame or list from which the variables in the formula should be
looked up.

* If no data frame or list is passed, then the parent frame is used.
* If no other arguments are passed, there are defaults that can be used.

**Lattice Behavior**

Lattice functions behave differently from base graphics functions in one critical way.

> * Base graphics functions plot data directly to the graphics device (screen, PDF file, etc.)
> * Lattice graphics functions return an object of class trellis
> * The print methods for lattice functions actually do the work of plotting the data on the graphics device.
> * Lattice functions return "plot objects" that can, in principle, be stored (but it’s usually better to just save the code + data).
> * On the command line, trellis objects are *auto-printed* so that it appears the function is plotting the data

**Summary**

> * Lattice plots are constructed with a single function call to a core lattice function (e.g. xyplot)
> * Aspects like margins and spacing are automatically handled and defaults are usually sufficient
> * The lattice system is ideal for creating conditioning plots where you examine the same kind of plot under many different conditions
> * Panel functions can be specified/customized to modify what is plotted in each of the plot panels


### [<kbd>Lesson 2</kbd>](#lesson-2) Plotting with ggplot2

**What is ggplot2?**

Grammer of Graphics

> “In brief, the grammar tells us that a statistical graphic is a mapping from data to aesthetic attributes (colour, shape, size) of geometric objects (points, lines, bars). The plot may also contain statistical transformations of the data and is drawn on a specific coordinate system” -- from ggplot2 book

Important points:

> * An implementation of The Grammar of Graphics by Leland Wilkinson
> * Written by Hadley Wickham (while he was a graduate student at Iowa State)
> * A “third” graphics system for R (along with base and lattice)
> * Grammar of graphics represents an abstraction of graphics ideas/objects
> * Think “verb”, “noun”, “adjective” for graphics
> * Allows for a “theory” of graphics on which to build new graphics and graphics objects
> * “Shorten the distance from mind to page”

**Plotting Systems in R**

<table>
	<tbody>
		<tr>
			<td>Base</td>
			<td>Lattice</td>
			<td>ggplot</td>
		</tr>
		<tr>
			<td>
				<ul>
					<li>&ldquo;Artist&rsquo;s palette&rdquo; model</li>
					<li>Start with blank canvas and build up from there</li>
					<li>Start with&nbsp;plot&nbsp;function (or similar)</li>
					<li>Use annotation functions to add/modify (text, lines, points, axis)</li>
					<li>Convenient, mirrors how we think of building plots and analyzing data</li>
					<li>Can&rsquo;t go back once plot has started (i.e. to adjust margins); need to plan in advance</li>
					<li>Difficult to &ldquo;translate&rdquo; to others once a new plot has been created (no graphical &ldquo;language&rdquo;)</li>
					<li>Plot is just a series of R commands</li>
				</ul>
			</td>
			<td>
				<ul>
					<li>Plots are created with a single function call (xyplot, bwplot,etc.)</li>
					<li>Most useful for conditioning types of plots: Looking at how&nbsp;<span id="MathJax-Element-1-Frame" class="MathJax"><span id="MathJax-Span-1" class="math"><span id="MathJax-Span-2" class="mrow"><span id="MathJax-Span-3" class="mi">y</span></span></span></span>&nbsp;changes with&nbsp;<span id="MathJax-Element-2-Frame" class="MathJax"><span id="MathJax-Span-4" class="math"><span id="MathJax-Span-5" class="mrow"><span id="MathJax-Span-6" class="mi">x</span></span></span></span>&nbsp;across levels of&nbsp;<span id="MathJax-Element-3-Frame" class="MathJax"><span id="MathJax-Span-7" class="math"><span id="MathJax-Span-8" class="mrow"><span id="MathJax-Span-9" class="mi">z</span></span></span></span></li>
					<li>Things like margins/spacing set automatically because entire plot is specified at once</li>
					<li>Good for putting many many plots on a screen</li>
					<li>Sometimes awkward to specify an entire plot in a single function call</li>
					<li>Annotation in plot is not intuitive</li>
					<li>Use of panel functions and subscripts difficult to wield and requires intense preparation</li>
					<li>Cannot &ldquo;add&rdquo; to the plot once it&rsquo;s created</li>
				</ul>
			</td>
			<td>
				<ul>
					<li>Split the difference between base and lattice</li>
					<li>Automatically deals with spacings, text, titles but also allows you to annotate by &ldquo;adding&rdquo;</li>
					<li>Superficial similarity to lattice but generally easier/more intuitive to use</li>
					<li>Default mode makes many choices for you (but you&nbsp;<em>can</em>&nbsp;customize!)</li>
				</ul>
			</td>
		</tr>
	</tbody>
</table>

**The basics: `qplot()`**

> * The qplot() function is the analog to plot() but with many built-in features
> * Syntax somewhere in between base/lattice
> * Produces very nice graphics, essentially publication ready (if you like the design)
> * Difficult to go against the grain/customize (don’t bother; use full ggplot2 power in that case)

**Basic Components of a ggplot2 Plot**

> * A data frame
> * aesthetic mappings: how data are mapped to color, size
> * geoms: geometric objects like points, lines, shapes.
> * facets: for conditional plots.
> * stats: statistical transformations like binning, quantiles, smoothing.
> * scales: what scale an aesthetic map uses (example: male = red, female = blue).
> * coordinate system
> * When building plots in ggplot2 (rather than using qplot) the “artist’s palette” model may be the closest analogy
> * Plots are built up in layers
> * Plot the data
>   * Overlay a summary
>   * Metadata and annotation

**Annotation**

> * Labels: xlab(), ylab(), labs(), ggtitle()
> * Each of the “geom” functions has options to modify
> * For things that only make sense globally, use theme()
>    * Example: theme(legend.position = "none")
> * Two standard appearance themes are included
>    * theme_gray(): The default theme (gray background)
>    * theme_bw(): More stark/plain

**Summary**

> * ggplot2 is very powerful and flexible if you learn the “grammar” and the various elements that can be tuned/modified
> * Many more types of plots can be made; explore and mess around with the package (references mentioned in Part 1 are useful)
